### 2016-09-23 review/response_to_reviewer.txt

Two reviewers raised the concern that active dendritic mechanisms
could qualitatively affect the claim of our study. We address their
concern in the following.

How could active dendritic mechanisms "qualitatively affect the claim
of our study" ? Our framework for single cell computation consists in
(i) evaluating analytically the somatic subthreshold membrane
potential fluctuations at the soma as a function of the presynaptic
quantities and (ii) convert those fluctuations into a spiking
probability thanks to a /firing response/ function determined /in
vitro/ in individual neuron thanks to somatic recordings.

Our framework thus needs to be tested against two phenomena: (I)
whether active dendritic mechanisms qualitatively affect the
relationship between presynaptic quantities and somatic membrane
potential fluctuations and (II) whether dendritic non-linearities
introduce a coupling between presynaptic activity and spike emission
that render the decomposition into the successive steps i & ii
misleading. Those phenomena will naturally appear to some extent and
will produce quantitative deviations with respect to our predictions
(that neglects dendritic non-linearities), the purpose of this study
is only to check for "qualitative deviations".

To investigate the impact of those phenomena, we included active
dendritic mechanisms (as provided by the model of [[citetext:Larkum2009]],
available on ModelDB, [[https://senselab.med.yale.edu/modeldb/showModel.cshtml?model=124043][link]]) into our framework for cellular
computation in the /fluctuation-driven/ regime. We included two of the
principal ingredients of active dendritic mechanisms into our model:
(i) NMDA currents (coupled to the AMPA synapses) and (ii) Ca^{2+}
spikes. We performed a numerical study on the /mean/ morphological
model of the main text (Table 1 and Figure 2,3), we spread those
mechanisms according to the picture provided by the study of
[[citetext:Larkum2009]], all glutamatergic synapses were replaced by the
AMPA+NMDA synapses and Ca^{2+} currents were included in the last two
branches to provide a /relay/ for far dendritic events (depicted in
the mean model drawing of [[fig:0a][Figure 1]]). Phenomenon (I) will be tested
against the inclusion of the NMDA mechanism and phenomenon (II) will
be tested against the inclusion of both NMDA and Ca^{2+} spikes. Those
two situations will be compared to the picture described in the main text.

#+NAME: fig:0a
#+CAPTION: *Including dendritic non-linearities (numerically) within our framework for single cell computation. On top of each panel, we present the recording of the membrane potential in the far distal dendrite (depicted with the upper electrode on the drawing) in the Control and Control+NMDA case. On the bottom, we present the membrane potential recorded at the soma (depicted with the lower electrode on the drawing) in four cases, the Control, the Control+NMDA case, the Control case with a HH model inserted in the soma and the Control+NMDA+Ca^{2+}spikes+HH(soma)*. *(A)* Response to baseline activity. \(\nu_e^p=\nu_e^d\)=0.2Hz, \(\nu_i^p=\nu_i^d\)=1.2Hz and s=0.05. *(B)* Response to baseline activity with the synchronous activation of 20 synapses every 500ms (marked by a triangle), note the impact of the NMDA mechanism under this stimulation type. *(C)* Response to an increase of proximal activity with respect to the baseline level (\(\nu_e^p\)=1.6Hz, \(\nu_i^p\)=10Hz). *(D)* Response to an increase in distally targetting presynaptic activity. *(E)* Response to an increase in unbalanced activity. *(F)* Response to an increase in 
[[./figs/traces.pdf]]

@@

show the their impact in the
fluctuation-driven regime, see [[fig:0b][Figure 2]].

#+NAME: fig:0b
#+CAPTION: *Effect of dendritic non-linearities in the subthreshold fluctuations and firing response in the fluctuation-dirven regime.* *(A)* 
[[./figs/active_sum_up.pdf]]

* Reviewer #1

\it Unfortunately, it is very difficult to grasp the essence and even more
the details of the study. \rm

1) \it Main statement; presentation. It starts with the title that claims
   that claims that “Heterogeneous Firing Responses Predict Diverse
   Couplings to Presynaptic Activity in mice Layer V Pyramidal
   Neurons”. The title refers to the phenomenology in Figure 5 showing
   that the output frequency goes up or down for different cells as a
   function of “increasing presynaptic quantity”, as the x-axis in
   that figure is label. But not clear how strong the increase is at
   the corresponding position, nor what of the presynaptic quantity
   is. For us (the writer and two other group members) it was neither
   possible to figure it out from the caption nor the text. \rm

    We agree that this way to represent data might not be
   straightforward. We improved the legend of Figure 4 to facilitate
   the comprehension. In particular, we have added an example to
   explicit the presence of "comodulations" of variables and hence the
   difficulty to represent the data as a function of one given
   variable. We believe that this way to present data is still the
   best one to emulate the comodulations that characterize neocortical
   dynamics (e.g. inhibition follows excitation) and to be able to
   represent the response to various forms of network activity
   (i.e. various comodulations) on individual cells.  The new legend
   reads:

   _The abscissa “increasing synaptic quantity” corresponds to the comodulations of the model variables shown in Figure 4A (same color code). As an example, the response to a “proximal activity increase” (blue curve) corresponds to a linear increase of and while keeping , and to their baseline level. See Figure 4A for the comodulations of the other types of increasing activity._

2) \it Main statement unclear. Unfortunately, it does not become clear
   what one has to understand of the “Diverse Couplings to the
   Presynaptic Activity” as expressed in the title. In the abstract it
   is stated: “We indeed found such a heterogeneous response, but
   trivially, the only significant contribution to the various firing
   response was the different levels of cellular excitability.” We
   guess that the “trivially” refers to the fact that only the
   cellular excitability has been modulated in the model. But then we
   don’t understand “A notable exception appeared for proximal
   activity: increasing proximal activity could either promote firing
   response in some cells or suppress it in some other cells whatever
   their individual excitability.” We conclude that there is still
   more than the individual excitability that determines the firing
   behavior, but this seems to contradict the previous statement. \rm

   We agree with the reviewer that the formulation "only [...] except
   [...]" might be misleading, we adapted the abstract to remove this
   source of confusion. Now it reads: "but trivially, the _main_
   significant contribution to the various firing response was the
   different levels of cellular excitability [...]".

@@

#+NAME: fig:1
#+CAPTION: *Changing the ratio of the proximal/distal separation.* *(A)* Configuration of the main text, the proximal part occupies the first 87.5% of the dendritic morphology. *(B)* Now the proximal part occupies only the first 12.5% of the dendritic morphology. This importantly reduces the number of available synapses (the morphology has only one branch in the first 12.5% part). Provided this loss of synapses is compensated by an increase of activity, there is _no qualitative change between the 12.5% and the 87.5% setting for the proximal activity: non-monotonic \(\sigma_V\) variations, strong decrease in \(\tau_V / \tau_m^0 \)_. On the other hand, this change would have important qualitative consequences for the "distal activity" regime, if the last 87.5% of the morphology are "distal", many synapses are electrotonically close (what is rather irrealistic for distal synapses), they will thus stronly interact through conductance shunting and thus produce fluctuations close to the "proximal activity" setting of the main text.
[[./figs/varying_prox_dist.png]]

3) \it Main statement; missing analysis. It is suggested that the
   heterogeneity would arise from sodium channels, but the paper is
   missing a joint presentation of experimental and modeling data
   showing some fit. Neither could we find a mathematical argument why
   the different phenomenological behavior shown in the cited Fig. 5
   should arise. It would be helpful to underscore the main statement
   by a simplified mathematical formulation in the main text. For
   instance, a simplified formula that gives the output firing rate as
   a function of the input variables would be helpful, including the
   dependencies on the sodium channel density. This sodium channel
   density is claimed to make the difference between the cell types
   according to the abstract: “various levels of density of sodium
   channels and sodium inactivation will control the response of
   individual neurons to proximally targeting presynaptic activity.”
   Again, we could not find a state-of-the-art explanation why this
   statement is true. \rm

    This point is precisely the focus of our previous communication
   [[cite:Zerlaut2016]]. Unfortunately there is, to the best of our knowledge, no
   closed-form solution that could give the firing rate response as a
   response to fluctuating input in a model including sodium
   inactivation and a sdium activation curve (hence the
   phenomenological approach formulated in ou previous
   communication). Nonetheless to give some intuition to the reader,
   we have added the following statement (line 316):
   _In our previous communication, theoretical modeling suggested that a high sensitivity to the speed of the fluctuations was enable by a high level of sodium inactivation (as only fast fluctuations allow to deinactivate sodium channels) and a high density of sodium channels (as it  produces a sharp spike initiation that enables to extract fast fluctuating input)._
   For further details, we believe the reader should refer to our
   previous publication.

4) \it No dendritic nonlinearities. One main characteristics of L5
   pyramidal neurons is the calcium spike in the apical tree, and more
   generally the various types of dendritic spikes. Non of them is
   considered in the model, and one conjectures that the various
   response properties in the experimental data arise from these
   various dendritic nonlinearities. Describing the firing properties
   of pyramidal neurons as a function of proximal and distal activity
   should take account of these dendritic nonlinearities. Otherwise,
   in fact, the statements cannot be compelling. \rm

   We adressed this concern in details in Part A.

5) \it Data extraction. It is not clear how the 4 cells have been selected
   out of the 27 cells. The 4 cells represent 4 cell types. How are
   these cell types specified? The paper does not give an answer on
   that. \rm

   The 4 cells represent 4 different cells out of the n=30 cells
   recorded in our previous study [[cite:Zerlaut2016]]. The criteria to
   choose those 4 cells was that they should somehow represent
   different behaviors while still being representative. See in Figure
   6B in [[citetext:Zerlaut2016]] how those 4 cells cover relatively well
   the domain of measured characteristics for the firing response,
   this is also the case in the present study, where they cover
   relatively well the domain of the predicted coupling to the various
   types of activities.

6) \it Figure 2 has two panels A & B, but the caption describes panels
   A–E. \rm

   Indeed, we apologize for this problem in the manuscript
   preparation. Figure 2 and Figure 3 have been mixed up...

7) \it Apart from that, the fit of the numerical simulations and the
   analytical estimate in B for sigma_V is biased. No argument is
   given why this bias is there. \rm

   From line 165 to line 182, we discuss the validity of our
   analytical treatment. We explicit the two hypothesis that underlie
   our estimation and detail their expected effect. We conclude by
   discussing the bias for \(\sigma)V\) from line 180 to 182: "Some
   discrepancies between the approximation and the simulations
   appeared, in particular one can see a 1mV shift in the standard
   deviation of the fluctuation (meaning that single events are
   underestimated in the analytical treatment, so that hypothesis 2 is
   the most problematic one)." (where hypothesis 2 has been explicited
   before).

8) \it The central claim is that distal and proximal inputs differently
   affects the output firing rates. Beside the fact that this
   statement by itself is trivial [...] \rm

   We also show that "distal and proximal inputs differently
   affects the output firing rates", but this is not our main
   statement and we agree with the reviewer that this could seen as
   "trivial". Our main statement claims that those stimulations affect
   individual neurons in diverse manners (in particular for the
   proximal stimulation: increasing their response for some neurons,
   decreasing it for some other neurons). We believe that this
   diversity of responses across neurons (to the same input) is not a
   so trivial phenomena.

9) \it [...] we wondered why the closer 87.5% of the dendritic branch is
   declared as “proximal” and the furthest 12.5% as the “distal”
   part. Is the result dependent on this choice, or could one have
   also chosen the closer 12.5% as “proximal” only? \rm

   Indeed, the _arbitrary ratio_ between proximal and distal input
   might need a more detailed discussion. Following the reviewer's
   suggestion, we reproduced the analysis with the convention
   "proximal as the closest 12.5% only". We show on [[fig:1][Figure 3]] that the
   results are not affected by this change, the qualitative properties
   of the fluctuations remain identical: strong decrease of the
   fluctuations time constant and non-monotonic shape of the standard
   deviation of the fluctuations. The key property leading to the
   properties of the "proximal activity" being that they correspond to
   synapses that are close to each others and to the soma (to produce
   a strong /shunting/ effect), this phenomena is even stronger under
   this setting. The reason for choosing a relatively large fraction
   for proximal synapses is to obtain a large number of
   synapses. Indeed, along the morphology, the electrotonic distance
   raises but the surface as well (because of the branches number, see
   drawing in Fig 1 in main text) so that the number of synapses
   strongly increases. Summing up, the parameter \(f_{prox}\)=87.5% is
   important to keep the distal part far enough electrotonically, for
   the proximal part there is no much difference between
   \(f_{prox}\)=87.5% and \(f_{prox}\)=12.5% (they all contain the
   very proximal synapses), and keeping \(f_{prox}\)=87.5% avoid to
   introduce an additional rescaling of synaptic frequences because of
   the reduced number of synapses (the procedure shown in [[fig:1][Figure 3]]).

@@

Spurious "s" in the sec "A simplified morphological model" - sWe arbitrarily ...
Where is excitability < V_thre^eff >_D defined?

\bf We apologize for the problems with the manuscript preparations and we thank the reviewer for pointing out those mistakes. \rm

* Reviewer #3

\it The key claim of this work is that recently identified L5 PCs firing
rate properties indicate how these cells couple to presynaptic
activity patterns (proximal, distal,...). The claim is based on i)
recently identified firing response functions (previous work of the
authors), ii) a theoretical model for passive dendritic integration
that is an extension of the single compartment model of
[[citetext:Kuhn2004]] to a passive tree with cable theory, and iii) an
identification of the passive tree of L5 PCs by fitting the input
impedance at the soma to experimental data.

### 2016-09-27 review/response_to_reviewer.txt

#+TITLE:  "Heterogeneous Firing Responses Predict Diverse Couplings to Presynaptic Activity in mice Layer V Pyramidal Neurons" \\ _\textsc{ Response to reviewers}_
#+AUTHOR: Y. Zerlaut & A. Destexhe

* Impact of active dendritic mechanisms

Two reviewers raised the concern that active dendritic mechanisms
(reviewed in [[citetext:Stuart2015]]) could qualitatively affect the claim
of our study. We address their concern in the following.

How could active dendritic mechanisms "qualitatively affect the claim
of our study" ? Our framework for single cell computation consists in
(*i*) evaluating analytically the somatic subthreshold membrane
potential fluctuations at the soma as a function of the presynaptic
quantities and (*ii*) convert those fluctuations into a spiking
probability thanks to a /firing response/ function determined /in
vitro/ in individual neuron thanks to somatic recordings.

Our framework thus needs to be tested against two phenomena: (*I*)
whether active dendritic mechanisms qualitatively affect the
relationship between presynaptic quantities and somatic membrane
potential fluctuations, i.e. checking the validity of step *i*, and
(*II*) whether dendritic non-linearities introduce a coupling between
presynaptic activity and spike emission that render the decomposition
into the successive steps *i* & *ii* misleading. Those phenomena will
naturally appear to some extent and will produce quantitative
deviations with respect to our predictions (that do not include
dendritic non-linearities, e.g. trivially, all cells should be more
excitable). As the main message of the paper does not rely on the
absolute values of the input-output functions but on the rather on the
relative behaviors from cell to cell that emerge from differences
across individual firing response functions, the purpose of this study
is only to investigate putative "qualitative deviations".

To investigate the impact of those phenomena, we included active
dendritic mechanisms (as provided by the model of [[citetext:Larkum2009]],
available on ModelDB on the following [[https://senselab.med.yale.edu/modeldb/showModel.cshtml?model=124043][link]]) into our framework for
cellular computation in the /fluctuation-driven/ regime. We included
two of the principal ingredients of non-linear dendritic mechanisms
into our model: NMDA currents (coupled to the AMPA synapse) and
Ca^{2+} currents. We performed a numerical study on the /mean/
morphological model of the main text (Table 1 and Figure 2,3), we
spread those mechanisms according to the picture provided by the study
of [[citetext:Larkum2009]], all glutamatergic synapses were replaced by
the AMPA+NMDA synapses and Ca^{2+} currents were included in the last
two branches to provide a /relay/ for far dendritic events (depicted
on the drawing of the mean model, see [[fig:0a][Figure 1]]). Phenomenon (*I*) will
be tested against the inclusion of the NMDA mechanism, we will examine
the properties of the membrane potential fluctuations at the soma,
with and without this mechanism. Phenomenon (*II*) will be tested
against the inclusion of both NMDA and Ca^{2+} currents: we will first
include numerically a spike emission model at the soma (Hodgkin-Huxley
like, Na^{+} and K^{+} currents) in the morphological model, then the
model will be able to produce the full input-output function (from
presynaptic quantities to spiking probability), we will then
investigate how NMDA and Ca^{2+} currents impacts the dependency on
the presynaptic variables for the full input-output function.

#+NAME: fig:0a
#+CAPTION: *Including dendritic non-linearities (numerically) within our framework for single cell computation. On top of each panel, we present the recording of the membrane potential in the far distal dendrite (depicted with the upper electrode on the drawing) in the Control and Control+NMDA case. On the bottom, we present the membrane potential recorded at the soma (depicted with the lower electrode on the drawing) in four cases, the Control, the Control+NMDA case, the Control case with a HH model inserted in the soma and the Control+NMDA+Ca^{2+}spikes+HH(soma)*. *(A)* Response to baseline activity. \(\nu_e^p=\nu_e^d\)=0.2Hz, \(\nu_i^p=\nu_i^d\)=1.2Hz and s=0.05. *(B)* Response to baseline activity with the synchronous activation of 20 synapses every 500ms (marked by a triangle), note the impact of the NMDA mechanism under this stimulation type. *(C)* Response to an increase of proximal activity with respect to the baseline level (\(\nu_e^p\)=1.6Hz, \(\nu_i^p\)=10Hz). *(D)* Response to an increase in distally targetting presynaptic activity (\(\nu_e^d\)=0.6Hz, \(\nu_i^d\)=3Hz). *(E)* Response to an increase in unbalanced activity (\(\nu_e^p\)=0.4Hz, \(\nu_i^p\)=2Hz,\(\nu_e^p\)=0.5Hz, \(\nu_i^p\)=1.7Hz). *(F)* Response to an increase in presynaptic synchrony (s=0.4).
[[./figs/traces.pdf]]

@@

#+NAME: fig:0b
#+CAPTION: *Effect of dendritic non-linearities in the subthreshold fluctuations and firing response in the fluctuation-dirven regime.* *(A)* 
[[./figs/active_sum_up.pdf]]

In [[fig:0a][Figure 1]], we represent the output of the numerical simulations of
the resulting models. We represent the subthreshold responses on top
plots, as well as the responses in presence of a spiking mechanism
(the HH model at the soma) on bottom plots. We show on [[fig:0a][Figure 1]]A,C-F
the response to the five protocols discussed within the
manuscript. Additionally, on [[fig:0a][Figure 1]]B, we added another type of
stimulus: the synchronous activation of 20 synapses on one distal
dendrite. The response to this last stimulus clearly exhibits the
presence of strong NMDA currents resembling those observed /in vivo/
[[cite:Larkum2009]]. We observe that, for the /baseline activity/, for the
/strong proximal activity/ and for the /strong distal activity/, there
does not seems to be a strong impact of the NMDA mechanism on the
subthreshold fluctuations (both at the soma and in the distal
dendrites): it exerts a slight depolarization but does not
qualitatively impacts the fluctuations. Neither the triggering of
action potentials seems to be affected (spikes are identically evoked
with and without NMDA and Ca^{2+} channels in presence of a somatic HH
model), see [[fig:0a][Figure 1]]A,C,D. On the other hand, for a high level of
unbalanced activity (where excitation globally overcomes inhibition),
we observe a strong depolarization induced by the NMDA mechanism, we
also observe that, consequently, the spiking probability seems
significantly increased, see [[fig:0a][Figure 1]]E. Also for a high level of
synchrony the fluctuations and spikes seems to be affected, we observe
the presence of small NMDA spikes and additional spikes are elicited
in the presence of NMDA and Ca^{2+} currents, see [[fig:0a][Figure 1]]E.

We then analyze with more details how those active mechanisms impacts
the dependency on the variables of the model. We thus reproduced the
analysis of Figure 3 in the main text. On [[fig:0b][Figure 2]]A, we report the
properties of the fluctuations as a function of the presynaptic
variables, with and without NMDA channels. We observe that NMDA
channels exert an additional small depolarization (increased
\(\mu_V\)), increase the standard deviation of the fluctuations
because they increase the amplitude of depolarizing events (increased
\(\sigma_V\)) and they slow down the fluctuations because they slowly
increase the low frequency content of post-synaptic depolarization
(increased \(\tau_V\)). We also observe, that, despite this small
shift in all quantities, the qualitative trend of all curves remain
identical. Only for high excitation a qualitative difference can be
observe (see bottom left plot of [[fig:0b][Figure 2]]A): increasing excitation
makes V_m fluctuations faster in the passive case whereas increasing
excitation makes fluctuations slower in the presence of NMDA currents
(decreasing vs increasing \(\tau_V\) curve respectively). We conclude
that phenomenon *I* has a very weak impact provided excitation does
not overcome inhibition and provided that synchrony remains in a
relatively low range (the range of our study: $s \leq 0.4\)).

On [[fig:0b][Figure 2]]B, we investigate how NMDA and Ca^{2+} currents shape the
dependency of the spiking probability on the presynaptic
variables. Trivially, we observe an increased spiking probability
because cells are more excitable (NMDA and Ca^{2+} currents are
depolarizing currents). Notably, when excitation does not balance
inhibition (\(\nu_e^p = 0.4\)Hz or \(\nu_i^p = 0\)Hz), the spiking
probability is greatly enhanced (almost doubled). For the other
variables, those effects are much more moderate. Here again, we
conclude that phenomenon *II* has a weak impact provided excitation
does not overcome inhibition and provided that synchrony remains in a
relatively low range (the range of our study: $s \leq 0.4\)).

As can be seen in [[fig:0a][Figure 1]]B, those results nonetheless apply only to
neocortical processing when effects are mediated by /weakly
synchronous and relatively balanced/ population dynamics, when single
cell response is driven by synchronous activation of multiple
synapses, dendritic nonlinearities have a crucial impact (consistent
with experimental evidences and the current view on dendritic
processing, reviewed in [[citetext:Stuart2015]]). We thus modified the
main text to briefly mention the results of the present analysis and
our conclusions. The main text now reads:

\bf The proposed theoretical framework for single-cell computation
nonetheless suffers from several weaknesses. [...]. Most importantly,
there is no active mechanisms in dendrites [[cite:Hausser2005]], it is
therefore a question how much those mechanisms could affect the
picture provided in our study. Preliminary numerical analysis
performed in presence of NMDA and Ca2+ currents [[cite:Larkum2009]],
showed that, provided excitation balances inhibition (a situation
where NMDA channels keeps a relatively low level of stationary
activation) and provided synchrony do not reach a too high level (so
that excitatory events have a weak probability to evoke NMDA and
Ca^{2+} spikes), the qualitative behavior of the cellular input-output
function remains identical. Our preliminary analysis thus suggest the
hypothesis that active dendritic mechanisms may not be crucial for
effects mediated by slow (T\(\geq\)20-30ms) variations of weakly synchronous
population dynamics while being critical for operations performed on
synchronous synaptic activations (T\(\leq\)5-10ms), consistent with the
current view on dendritic processing [[cite:Stuart2015]]. Future work
should address this hypothesis in details. \rm

Because we believe that those results constitute only preliminary data
and that this problem should be the focus of a much more detailed
analysis, we do not wish to include it in the main text or
supplementary material. If the reviewer think that this is important
for our study, we can merge [[fig:0a][Figure 1]] and [[fig:0b][Figure 2]] and produce a
supplementary figure whose legend will be adapted from this section
and the current legends.

@@

\it Unfortunately, it is very difficult to grasp the essence and even more
the details of the study. \rm

1) \it Main statement; presentation. It starts with the title that claims
   that claims that “Heterogeneous Firing Responses Predict Diverse
   Couplings to Presynaptic Activity in mice Layer V Pyramidal
   Neurons”. The title refers to the phenomenology in Figure 5 showing
   that the output frequency goes up or down for different cells as a
   function of “increasing presynaptic quantity”, as the x-axis in
   that figure is label. But not clear how strong the increase is at
   the corresponding position, nor what of the presynaptic quantity
   is. For us (the writer and two other group members) it was neither
   possible to figure it out from the caption nor the text. \rm

    We agree that this way to represent data might not be
   straightforward. We improved the legend of Figure 4 to facilitate
   the comprehension. In particular, we have added an example to
   explicit the presence of "comodulations" of variables and hence the
   difficulty to represent the data as a function of one given
   variable. We believe that this way to present data is still the
   best one to emulate the comodulations that characterize neocortical
   dynamics (e.g. inhibition follows excitation) and to be able to
   represent the response to various forms of network activity
   (i.e. various comodulations) on individual cells.  The new legend
   reads:

   \bf The abscissa “increasing synaptic quantity” corresponds to the
   comodulations of the model variables shown in Figure 4A (same color
   code). As an example, the response to a “proximal activity
   increase” (blue curve) corresponds to a linear increase of
   \(\nu_e^{prox}\) and \(\nu_i^{prox}\) while keeping
   \(\nu_e^{dist}\), \(\nu_i^{dist}\) and \(s\) to their baseline
   level. See Figure 4A for the comodulations of the other types of
   increasing activity. \rm

@@

2) \it Main statement unclear. Unfortunately, it does not become clear
   what one has to understand of the “Diverse Couplings to the
   Presynaptic Activity” as expressed in the title. In the abstract it
   is stated: “We indeed found such a heterogeneous response, but
   trivially, the only significant contribution to the various firing
   response was the different levels of cellular excitability.” We
   guess that the “trivially” refers to the fact that only the
   cellular excitability has been modulated in the model. But then we
   don’t understand “A notable exception appeared for proximal
   activity: increasing proximal activity could either promote firing
   response in some cells or suppress it in some other cells whatever
   their individual excitability.” We conclude that there is still
   more than the individual excitability that determines the firing
   behavior, but this seems to contradict the previous statement. \rm

   We agree with the reviewer that the formulation "only [...] except
   [...]" might be misleading, we adapted the abstract to remove this
   source of confusion. Now it reads: \bf "but trivially, the _main_
   significant contribution to the various firing response was the
   different levels of cellular excitability [...]". \rm

@@

3) \it Main statement; missing analysis. It is suggested that the
   heterogeneity would arise from sodium channels, but the paper is
   missing a joint presentation of experimental and modeling data
   showing some fit. Neither could we find a mathematical argument why
   the different phenomenological behavior shown in the cited Fig. 5
   should arise. It would be helpful to underscore the main statement
   by a simplified mathematical formulation in the main text. For
   instance, a simplified formula that gives the output firing rate as
   a function of the input variables would be helpful, including the
   dependencies on the sodium channel density. This sodium channel
   density is claimed to make the difference between the cell types
   according to the abstract: “various levels of density of sodium
   channels and sodium inactivation will control the response of
   individual neurons to proximally targeting presynaptic activity.”
   Again, we could not find a state-of-the-art explanation why this
   statement is true. \rm

   This point is precisely the focus of our previous communication
   [[cite:Zerlaut2016]]. Unfortunately there is, to the best of our
   knowledge, no closed-form solution that could give the firing rate
   response as a response to fluctuating input in a model including
   sodium inactivation and a sodium activation curve (hence the
   phenomenological approach formulated in our previous
   communication). Nonetheless to give some intuition to the reader,
   we have added the following statement (line 316): \bf In our
   previous communication, theoretical modeling suggested that a high
   sensitivity to the speed of the fluctuations was enable by a high
   level of sodium inactivation (as only fast fluctuations allow to
   deinactivate sodium channels) and a high density of sodium channels
   (as it produces a sharp spike initiation that enables to extract
   fast fluctuating input). \rm For further details, we believe the
   reader should refer to our previous publication.

@@

4) \it No dendritic nonlinearities. One main characteristics of L5
   pyramidal neurons is the calcium spike in the apical tree, and more
   generally the various types of dendritic spikes. Non of them is
   considered in the model, and one conjectures that the various
   response properties in the experimental data arise from these
   various dendritic nonlinearities. Describing the firing properties
   of pyramidal neurons as a function of proximal and distal activity
   should take account of these dendritic nonlinearities. Otherwise,
   in fact, the statements cannot be compelling. \rm

   We adressed this concern in details in Part A.

5) \it Data extraction. It is not clear how the 4 cells have been selected
   out of the 27 cells. The 4 cells represent 4 cell types. How are
   these cell types specified? The paper does not give an answer on
   that. \rm

   The 4 cells represent 4 different cells out of the n=30 cells
   recorded in our previous study [[cite:Zerlaut2016]]. The criteria to
   choose those 4 cells was that they should somehow represent
   different behaviors while still being representative. See in Figure
   6B in [[citetext:Zerlaut2016]] how those 4 cells cover relatively well
   the domain of measured characteristics for the firing response,
   this is also the case in the present study, where they cover
   relatively well the domain of the predicted coupling to the various
   types of activities, see Figure 6 in main text. We updated the
   manuscript to highlight this point (line 250): \bf We show on
   Figure 5 the response of four cells (representative of different
   firing response behaviors, see Figure 5 and 6B in
   [[citetext:Zerlaut2016]]) to the different types of presynaptic
   activity described in the previous section. The input-output
   relationships show qualitative and quantitative differences, [...]
   \rm

@@

6) \it Figure 2 has two panels A & B, but the caption describes panels
   A–E. \rm

   Indeed, we apologize for this problem in the manuscript
   preparation. Figure 2 and Figure 3 have been mixed up, this is now
   corrected.

@@

9) \it [...] we wondered why the closer 87.5% of the dendritic branch is
   declared as “proximal” and the furthest 12.5% as the “distal”
   part. Is the result dependent on this choice, or could one have
   also chosen the closer 12.5% as “proximal” only? \rm

   Indeed, the _arbitrary ratio_ between proximal and distal input
   might need a more detailed discussion. Following the reviewer's
   suggestion, we reproduced the analysis with the convention
   "proximal as the closest 12.5% only". We show on [[fig:1][Figure 3]] that the
   results are not affected by this change, the qualitative properties
   of the fluctuations remain identical: strong decrease of the
   fluctuations time constant and non-monotonic shape of the standard
   deviation of the fluctuations. The key property leading to the
   properties of the "proximal activity" being that they correspond to
   synapses that are close to each others and to the soma (to produce
   a strong /shunting/ effect), this phenomena is even stronger under
   this setting. The reason for choosing a relatively large fraction
   for proximal synapses is to obtain a large number of
   synapses. Indeed, along the morphology, the electrotonic distance
   raises but the surface as well (because of the branches number, see
   drawing in Fig 1 in main text) so that the number of synapses
   strongly increases. Summing up, the parameter \(f_{prox}\)=87.5% is
   important to keep the distal part far enough electrotonically, for
   the proximal part there is no much difference between
   \(f_{prox}\)=87.5% and \(f_{prox}\)=12.5% (they all contain the
   very proximal synapses), and keeping \(f_{prox}\)=87.5% avoid to
   introduce an additional rescaling of synaptic frequences because of
   the reduced number of synapses (the procedure shown in [[fig:1][Figure 3]]).

* Reviewer #2

 \it This manuscript shows a nice mixture of experiment and
 theory. The latter is based upon sensible use of the cable modelling
 framework of Rall combined with stochastic models of synaptic input
 (admittedly of a particular type but definitely fit for purpose). \rm

1) \it Unfortunately some of the work (for the firing response function)
   is buried in supplementary material - and this has not been made
   available to me (S1).\rm

   That is indeed unfortunate. For us, the permanent link at the end
   of the submission pdf (page 32/32) was functional and actually gave
   the supplementary material. Maybe the reviewer could ask the editor
   for this material if this is still not available in this
   resubmission.

2) \it It would be especially informative to hear more about the author's
   perspective on how synchrony controls rate. Nonetheless given the
   quality and novelty of the main manuscript I suggest to put this
   issue aside. \rm
